---
title: "ESM244 HW2"
author: "Sara Orofino"
date: "2/6/2019"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}

#load packages
library(tidyverse)
library(RColorBrewer)
library(tinytex)


# read in data 
fish <- read_csv("fish_catch_clean.csv")
cetacean <- read_csv("captive_cetacean_regions.csv")
```


#{.tabset}
##Task 1

###Data Visualization - Cetaceans in Captivity

```{r wrangle1, message=FALSE, warning=FALSE}

# Organize origin_location by US region according to US Census Bureau:
# West - Marineland of the Pacific, SeaWorld San Diego, Six Flags Discovery Kingdom, Dolphin Experience, Dolphin Quest Hawaii, Kahala Hilton Oahu, Sea Life Park, Mirage Dolphin Habitat
## West includes Hawaii, California, Nevada
# South - Discovery Cove, Dolphin Connection, Dolphin Research Center, Dolphin Plus Bayside, Dolphin Plus Oceanside, Gulf Wolrd Marine Park, Gulfarium, Marineland Florida, Miami Sequarium, Theater of the Sea, SeaWorld San Antonio, Georgia Aquarium,  
## South includes Florida, Texas, Georgia, 
# Midwest - Minnesota Zoo, Indianapolis Zoo, Brookfield Zoo, Shedd Aquarium, SeaWorld of Ohio, SeaWorld Aurora 
## Midwest includes Minnesota, Indiana, Ohio and Illinois 
# Northeast - National Aquarium, New England Aquarium 
## Northeast - Maryland and Massachusetts 
# Military - U.S. Navy and U.S. Navy MMP 
# International - Dolphin Quest Bermuda, Marineland Antibes, Marineland Canada, Sealand of the Pacific

acq_data <- cetacean %>% 
    filter(acquisition =="Born",
           birthYear != "NA",
           birthYear >= 1990,
           birthYear <= 2015) %>% 
  select(species, birthYear, us_regions) %>% 
  group_by(us_regions, birthYear) %>% 
  count() 

```


```{r}
#stacked column graph 

ggplot(acq_data, aes(x = birthYear, y = n , group = us_regions, fill = us_regions))+
  geom_bar(stat = "identity") + 
  scale_y_continuous(expand = c(0,0), limits = c(0,31), breaks = seq(0,30, by = 5)) + 
  scale_x_discrete(expand = c(0,0)) + 
  scale_fill_discrete(labels = c("International", "U.S. Midwest", "U.S. Military",
                                 "U.S. Northeast", "U.S. South", "Unknown", "U.S. West")) +
  theme_classic()+
  guides(fill=guide_legend(title="Region of Origin")) +
  labs(y = "Number of Cetaceans",
       x  = "Year") + 
  coord_flip()

```

**Figure 1. Cetaceans Born in Captivity by Region 1990-2015.** The graph depicts the number of cetaceans born into captivity each year from 1990-2015. Data only includes cetaceans whose birth years are known. Each color indicates the proportion of cetaceans born in each region for the given year. Regions distinctions in the United States are based on census regions from the United States Census Bureau.  
Data: Amber Thomas, The Pudding  
Sources: U.S. National Marine Mammal Inventory June 2015, Ceta-Base May 2017. 



##Task 2 

###Parameter Estimation - Wild Fish Catch 

a. Exploratory Graph of wild catch over time:
  
```{r exploratory_graph, echo=FALSE}

colnames(fish) <- c("year", "wild", "farmed", "total") 

fish_growth <- fish %>% 
  mutate(time = 0:62) %>% 
  select(time, wild)
  
  
  
ggplot(fish_growth, aes(x=time, y=wild)) +
  geom_point() + 
  scale_y_continuous(limits = c(0,100), breaks = seq(0,100, by =20))


```

b. What type of relationship describes the trend?  

The graph depicts logistic growth.  

Mathematically:    

$N_t = \left(\frac{A}{1 + \beta e^{-rt}}\right)$  

Initial Parameter estimates:  

Find estimates for logistic growth (A, B and r):

Carrying Capacity (A) from the exploratory graph above:  
A = 90 million tons

If $N_0 = 17$ million tons estimate $\beta$:      
$\beta = \frac{(K - N_0)}{N_0}$   
$\beta = \frac{(90-17)}{17} = 4.294$  

```{r estimate_b}

Bestimate <- (90-17)/17
```

Estimate r:

$r = 0.06648$   

```{r estimate_r}

lag_phase <- fish_growth %>% 
  filter(time < 19, time > 14)
  
ggplot(lag_phase, aes(x= time, y = log(wild))) +
  geom_point()
  
lm(log(wild) ~ time, data = lag_phase)

#r=0.006648
```

c. Model Parameters using Nonlinear Least Squares:   

```{r}

fish_fit <- nls(wild ~ A/(1+B*exp(-r*time)),
                start = list(A = 90, B = 4.29, r = 0.06648),
                data = fish_growth, 
                trace = TRUE)
fish_fit
# store coefficients for final graph:

A <- coef(fish_fit)[1]
B <- coef(fish_fit)[2]
r <- coef(fish_fit)[3]
```
Carrying Capacity (A):  
$A = 100.28$ million tons   

($\beta$):  
$\beta =  4.316$   

Growth Rate (r):  
$r = 0.06989$    

d. Finalized Graph:  
  
```{r final_graph, message=FALSE, warning=FALSE}

# Create a new sequence of time values to make predictions and model it: 

time_seq <- seq(0,70, length=100) 

fish_pred <- A/(1+B*exp(-r*time_seq))

# Bind with time sequence data:

pred_df <- data.frame(time_seq, fish_pred)

#Final graph: 

ggplot(fish_growth, aes(x = time, y = wild)) +
  geom_point(color = "slategray", size = 2, alpha = 0.7) +
  geom_line(data = pred_df, aes(x = time_seq, y = fish_pred), color = "midnightblue", size = 1) + 
  scale_x_continuous(expand = c(0,0.5), limits = c(0,70), breaks = seq(0,70, by = 10)) +
  scale_y_continuous(expand = c(0,0), limits = c(0,100)) +
  labs(x = "Time \n(Time 0 = 1950)", y = "Wild Fish Catch (Million Tons)") +
  theme_bw() 

```

##Task 3 

###Boostrapped Confidence Interval for Proportions 



##Task 4

###R-Studio Conference Talks 
